---
title: "Optional_assignment"
author: "Marcelladane"
date: "27 July 2017"
output: html_document
---

**Instructions:**

To practice the plotting techniques you have learned so far, you will be making a graphic that explores relationships between variables. This practice is useful since we will later cover creating reproducible graphics in this class. You will be looking at a subset of a United States medical expenditures dataset with information on costs for different medical conditions and in different areas of the country.

**You should do the following:**

**1** Make a plot that answers the question: *what is the relationship between mean covered charges (Average.Covered.Charges) and mean total payments (Average.Total.Payments) in New York?*

**2** Make a plot (possibly multi-panel) that answers the question: *how does the relationship between mean covered charges (Average.Covered.Charges) and mean total payments (Average.Total.Payments) vary by medical condition (DRG.Definition) and the state in which care was received (Provider.State)?*

First you need to read the CSV file.
```{r}

packages <- c("data.table", "reshape2", "ggplot2")
sapply(packages, require, character.only = TRUE, quietly = TRUE)

data <- read.csv(file = "payment.csv", header = TRUE, sep = ",")
head(data)
```

For the first question you must subset just the data from NY

You can use *grepl* and inform that the data is inside the variable Provider.State

Then you can make a cleaner table (decrease unnecessary size) with just *Average.Covered.Charges* and *Average.Total.Payments*
```{r}
subNY <- grepl("NY", data$Provider.State, ignore.case = TRUE)
NY_data <- data[subNY,]
NY_data_selected <- NY_data[,c("Average.Covered.Charges" , "Average.Total.Payments")]
```

Now that you have a table with just the data you need


You can plot it

Plot details:

Set the X and Y variable

add titles using xlab, ylab, main

Use alpha to introduce some transparency in the dots in the plot, its easier to visualize

Set pch for whatever shape you want, I wanted 11 (a star)

Since the graph is relationship, I added abline lm, to create a linear regression for the relation of Av.Payments and Av.Charges
```{r}
pdf("plot1.pdf")
plot(NY_data_selected$Average.Covered.Charges, NY_data_selected$Average.Total.Payments, xlab = "Av. Charges ($)", ylab = "Av. Total Payments ($)",
     main = "Relationship between Average in Charges and Payments in NY", col = adjustcolor("green", alpha = 0.4), pch = 11)
abline(lm(NY_data_selected$Average.Total.Payments ~ NY_data_selected$Average.Covered.Charges), col = "red", lwd = 1)
dev.off()
```

