---
title: "Optional_assignment_p2"
author: "Marcelladane"
date: "27 July 2017"
output: html_document
---


**Instructions:**

To practice the plotting techniques you have learned so far, you will be making a graphic that explores relationships between variables. This practice is useful since we will later cover creating reproducible graphics in this class. You will be looking at a subset of a United States medical expenditures dataset with information on costs for different medical conditions and in different areas of the country.

**You should do the following:**

**1** Make a plot that answers the question: *what is the relationship between mean covered charges (Average.Covered.Charges) and mean total payments (Average.Total.Payments) in New York?*

**2** Make a plot (possibly multi-panel) that answers the question: *how does the relationship between mean covered charges (Average.Covered.Charges) and mean total payments (Average.Total.Payments) vary by medical condition (DRG.Definition) and the state in which care was received (Provider.State)?*

First you need to read the CSV file.
```{r}

packages <- c("data.table", "reshape2", "ggplot2")
sapply(packages, require, character.only = TRUE, quietly = TRUE)

data <- read.csv(file = "payment.csv", header = TRUE, sep = ",")
head(data)
```

The first thing is to check the levels of the variable DRG
```{r}
levels(data$DRG.Definition)
```

As you can see the names are too big and will make a mess in the graphic, so I will transform them to be just the number and the inicials

| Real name                                      | Abreviation   |
| ---------------------------------------------- |:-------------:|
| SIMPLE PNEUMONIA & PLEURISY W CC               | SP            |
| HEART FAILURE & SHOCK W CC                     | HF            |
| ESOPHAGITIS, GASTROENT & MISC D. DIS. W/O MCC  | DD            |
| MISC DISORDERS OF NUTRITION,METABOLISM ...     | DD            |
| KIDNEY & URINARY TRACT INFECTIONS W/O MCC      | UI            |
| SEPTICEMIA OR SEVERE SEPSIS W/O MV 96+...      | SS            |

```{r}
levels(data$DRG.Definition) <- c("194 - SP", "292 - HF", "392 - DD", "641 - ND", "690 - UI", "871 - SS")
```

Now I will subset again just the part of the data I want to use *Average.Covered.Charges*, *Average.Total.Payments*, *Provider.State* and *DRG.Definition*
```{r}
data_selected <- data[,c("DRG.Definition" , "Average.Covered.Charges" , "Average.Total.Payments", "Provider.State")]
levels(data$Provider.State)
head(data_selected)
```

Now I must to set key words for the plot, since the plot will be divided in STATE + CONDITION

Use *unique* to extract unique elements
```{r}
State <- unique(data$Provider.State)
Condition <- unique(data$DRG.Definition)
```

*Now for the plot*

First I need to make a plot that will hold my 6 conditions and 6 States

Increase margins of the plot to fit all 36 plots

Since they are 36 plots, I will use a *for* loop to make all of them. 

Use 2 dummy variables to call the key words "State" and "Condition" selected before

use with and subset that Provider.State is "a" and DRG.Definition is "b"

Then you can actually call the plot

You will need to set limits for y and x axis, otherwise they will be different for each state and will be harder to visualize real differences

Wanted to plot in magenta and with a transparency of 0.3 with dollar signs as symbol (just because the subject is cost)

Add a abline for linear regression (lm) to show relationship of Total payment and charges

Since this plot is multiple, set some main titles with *mtext*
```{r}
pdf("plot2.pdf")

par(mfrow = c(6,6), oma = c(4,4,4,2), mar = rep(2,4))

for (a in State) {
  for (b in Condition){
    with(subset(data_selected, Provider.State == a & DRG.Definition == b),
         plot(Average.Covered.Charges, Average.Total.Payments, 
         main = paste(a,b),
         ylim = range(data_selected$Average.Total.Payments),
         xlim = range(data_selected$Average.Covered.Charges),
         col = adjustcolor("magenta", alpha = 0.3), pch = 35)
    )
    abline(lm(Average.Total.Payments~Average.Covered.Charges,
           subset(data_selected, Provider.State == a & DRG.Definition == b)), 
           col = "red")
  }
}

mtext("Av. Covered Charges ($)", side = 1, outer = TRUE, line = 1)
mtext("Av. Total Payments ($)", side = 2, outer = TRUE, line = 1)
mtext("Relationship between Covered Charges and Total Payments \n (Divided by State and Condition clustes)", side = 3, outer = TRUE, font = 2,
      line = 1)

dev.off()
```


